---
title: "A study of optimization algorithms on a breast cancer diagnosis dataset"
author: "Xinlei Chen, Guojing Wu and Yujing Yao"
abstract: "This report discusses a study of optimization algorithms on breast cancer diagnosis dataset. Our goal is to build a predictive model based on logistic regression to facilicate cancer diagnosis, and we compared estimation method including Newton Raphson, Gradient Decent and Lasso. Our result shows that....."
thanks:
keywords: "Gradient Decent; Lasso Logistic; Newton Raphson; Optimization"
date: "March 14, 2019"
output:
    pdf_document:
    highlight: default
    number_sections: true
    citation_package:
    keep_tex: false
    fig_caption: true
    latex_engine: pdflatex
fontsize: 10pt
geometry: margin=1in
bibliography:
biblio-style:
header-includes:
- \usepackage{indentfirst}
- \usepackage{graphicx}
- \usepackage{geometry}
- \usepackage{subfigure}
- \usepackage{amsmath}
- \usepackage{listings}
- \usepackage{tikz}
- \usetikzlibrary{matrix}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, comment = "")
library(tidyverse)
library(Hmisc)#for errvar
library(ggcorrplot) # for correlation heatmap
library(ggpubr) # common legend
library(matrixcalc) #is.negative.def
library(kableExtra) # table
set.seed(99999)
options(knitr.table.format = "latex")
```

## Background

The goal of the exerise is to build a predictive model based on logistic regression to facilicate cancer diagnosis.

### Dataset

The data \textit{breast-cancer.csv} have 569 row and 33 columns. The first column \textbf{ID} lables individual breast tissue images; The second column \textbf{Diagnonsis} indentifies if the image is coming from cancer tissue or benign cases (M=malignant, B = benign). There are 357  benign and 212  malignant cases. The other 30 columns correspond to mean, standard deviation and the largest values (points on the tails) of the distributions of the following 10 features computed for the cellnuclei.
\begin{itemize}
\item radius (mean of distances from center to points on the perimeter)
\item texture (standard deviation of gray-scale values)
\item perimeter
\item area
\item smoothness (local variation in radius lengths)
\item compactness ($perimeter^2/area$ - 1.0)
\item concavity (severity of concave portions of the contour)
\item concave points (number of concave portions of the contour)
\item symmetry
\item fractal dimension ("coastline approximation" - 1)
\end{itemize}


```{r}
###### data manipulation
#setwd("/Users/yujingyao/Desktop/")
mydata <- read.csv("breast-cancer-1.csv") 
n <- dim(mydata)[1]
p <- dim(mydata)[2]
list <- c(3:(p - 1))
namesX <- names(mydata)[-c(1,2,p)]
# standardize
dataX <- do.call(cbind, lapply(list, function(x) (mydata[,x] - mean(mydata[,x]))/sd(mydata[,x]))) 
# design matrix
X <- data.frame(dataX) %>% 
  mutate(., intercept = 1)
# response
resp <- as.vector(ifelse(mydata[,2] == "M", 1, 0))
```

### Variable selection

2. check multicolinearity -one plot

**Figure 1: multicolinearity plot of the dataset**

```{r}
###### plot to check collinearity
colnames(dataX) <-namesX
colinearity.plot <- function(data){
  data.frame(data) %>% 
  select(starts_with("radius"), 
         starts_with("texture"), 
         starts_with("perimeter"), 
         starts_with("area"), 
         starts_with("smooth"), 
         starts_with("compact"), 
         starts_with("concavity"), 
         starts_with("concave"), 
         starts_with("symmetry"), 
         starts_with("fractal")) %>% 
  cor() %>% 
  ggcorrplot(.,ggtheme = ggplot2::theme_gray,
             colors = c("#6D9EC1", "white", "#E46726"),
             tl.cex = 6)}
g1 <- colinearity.plot(dataX)

###### variable selection
eig <- eigen(cor(dataX))$values
# found values very close to 0, multicolinearity exists
# function: find the maximum correlation between i and j
max.corr <- function(data){
  len = dim(data)[2]
  a <- 0.5
  for (i in 1:(len-1)) {
    for (j in (i+1):len) {
      if(abs(cor(data[,i],data[,j]))> a) a <- cor(data[,i],data[,j])
    }
  }
  return(round(a,3))
}
# function: update dataset according to several rules:eigenvalues and corr
selection <- function(data,eigen.tol,corr.tol){
  while (min(eigen(cor(data))$values) <= eigen.tol & max.corr(data) >= corr.tol){
       temp <- data
       data <- temp[,-(which(round(abs(cor(temp)),3) == max.corr(temp),arr.ind = TRUE)[1,1])]
      }
  return(data)}
newdataX <- selection(dataX, eigen.tol=1e-2,corr.tol = 0.8)
g2 <- colinearity.plot(newdataX)
eigpost <- eigen(cor(newdataX))$values
# look at the difference of the deleted colums
delnames <- setdiff(colnames(dataX),colnames(newdataX))
ggarrange(g1,g2, ncol=2, nrow=1, common.legend = TRUE, legend="bottom")
```

```{r}

```


3. report final input of model - descripive mean(sd)



## Method

### Logistic Regression

Let $y$ be the vector $n$ response random variable, $X$ denote the $n\times p$ design matrix(let$X_{i}$ denote the $i$th row) and $\beta$ denote the $p\times 1$ coefficient.
The likelihood of logistic regression is:
$$L(\beta; X, y) = \prod_{i=1}^n \{(\frac{\exp(X_{i}\beta)}{1+\exp(X_{i}\beta)})^{y_i}(\frac{1}{1+\exp(X_{i}\beta)})^{1-y_i}\}$$
Maximizing the likelihood is equivalent to maximizing the log likelihood:
$$
\begin{aligned}
f(\beta) 
& = \sum_{i=1}^n \{y_i(X_{i}\beta)-\log(1+\exp(X_{i}\beta))\}\\
& = <X\beta, Y> - \sum_{i=1}^n\log(1+\exp(X_{i}\beta))
\end{aligned}
$$
Let $p$, a vector of $n$ denote $p=\frac{\exp(X\beta)}{1+\exp(X\beta)}$. 
The gradient of this function is:
$$\nabla f(\beta) = X^T(y-p)$$
The Hessian is given by:
$$\nabla^2 f(\beta) = -X^T W X$$ where $W = diag(p_1(1-p_1),p_2(1-p_2),\cdots,p_n(1-p_n))$
Hessian matrix is negative definite, well behaved.

```{r}
###### 0.logistic regression
logfit.0 <- glm(resp~dataX, family = binomial(link = "logit"))
# algorithm didn't converge without delete colinearity
logdata <- cbind.data.frame(resp,newdataX)
logfit.1 <- glm(resp~., family = binomial(link = "logit"),data = logdata)
logit.beta <- coef(logfit.1)
```


### Newton Raphson
```{r}
##### 1. classical newton raphson
newX <- data.frame(newdataX) %>%
  mutate(., intercept = 1)
newdat <- list(y = resp, X = as.matrix(newX))
# function: calcualte loglik, gradient, hessian
logisticstuff <- function(dat, betavec){ 
  u <- dat$X %*% betavec
  expu <- exp(u) 
  loglik <- t(u) %*% dat$y - sum((log(1 + expu))) # Log-likelihood at betavec
  prob <- expu / (1 + expu) # P(Y_i=1|x_i)
  grad <- t(dat$X) %*% (dat$y - prob)
  Hess <- -t(dat$X) %*% diag(as.vector(prob*(1 - prob))) %*% dat$X # Hessian at betavec
  return(list(loglik = loglik, grad = grad, Hess = Hess))}
NewtonRaphson <- function(dat, func, start, tol=1e-10, maxiter =200){ 
  i <- 0 
  cur <- start 
  stuff <- func(dat, cur) 
  res <- c(0, stuff$loglik, cur)
  prevloglik <- -Inf 
  while (i < maxiter && abs(stuff$loglik - prevloglik) > tol && stuff$loglik > -Inf) {
    i <- i + 1 
    prevloglik <- stuff$loglik 
    prev <- cur 
    cur <- prev - solve(stuff$Hess) %*% stuff$grad
    # # step halving
    #     lambda <- 1
    #     curtemp <- prev - solve(stuff$Hess) %*% stuff$grad 
    #     while(func(dat, curtemp)$loglik <= func(dat,prev)$loglik){
    #       lambda <- lambda*(1/2)
    #       curtemp <- prev - solve(stuff$Hess) %*% stuff$grad
    #     }
    #     cur <- curtemp
    # # step halving
    stuff <- func(dat, cur) 
    res <- rbind(res, c(i, stuff$loglik, cur))}
  return(res)
}
newres <- NewtonRaphson(newdat,logisticstuff, start = rep(0, dim(newdat$X)[2]))
# check convergence
check <- tail(newres)[,1:2]
newton.beta <- newres[nrow(newres),3:dim(newres)[2]]
```


### Gradient Decent
```{r}
##### 2. gradient descent
gradient <- function(dat, func, start, tol=1e-10,maxiter =200){ 
  i <- 0 
  cur <- start 
  pp <- length(start)
  stuff <- func(dat, cur) 
  hess <- t(dat$X)%*%(dat$X)# double check
  res <- c(0, stuff$loglik, cur)
  prevloglik <- -Inf
  while (i < maxiter && abs(stuff$loglik - prevloglik) > tol && stuff$loglik > -Inf) { 
    i <- i + 1 
    prevloglik <- stuff$loglik 
    prev <- cur 
    cur <- prev + solve(hess) %*% (stuff$grad)
    stuff <- func(dat, cur) 
    res <- rbind(res, c(i, stuff$loglik, cur))} 
  return(res)
}
gradres <- gradient(newdat, logisticstuff, start = rep(0, dim(newdat$X)[2]),maxiter =200)
# check convergence
check <- tail(gradres)[,1:2]
grad.beta <- gradres[nrow(gradres),3:dim(gradres)[2]]
```

### Lasso
problem 3&4 
```{r}
##### 3. coordinate-wise logistic lasso
sfun <- function(beta,lambda) sign(beta) * max((abs(beta) - lambda), 0)
coordinatelasso <- function(lambda, dat, start, tol=1e-10, maxiter = 200) {
  i <- 0 
  pp <- length(start)
  n <- length(dat$y)
  betavec <- start 
  u <- (dat$X) %*% betavec
  expu <- exp(u) 
  prob <- expu/(expu+1)
  # weighted updates
  w <- prob*(1-prob)
  z <- u + (dat$y-prob)/w
  loglik <- t(u) %*% dat$y - sum(t(log(1 + expu))) - sum(lambda * abs(betavec))
  # quardratic approximation to the log likelihood
  # loglik <- 1/2*(1/n)*sum(w*(z-u)^2) - sum(lambda * abs(betavec)) 
  res <- c(0, loglik, betavec)
  prevloglik <- -Inf # To make sure it iterates 
  while(i < maxiter && abs(loglik - prevloglik) > tol && loglik > -Inf) {
    i <- i + 1 
    prevloglik <- loglik
    prev <- betavec
    for (j in 1:pp){
      ynoj <- dat$X[,-j] %*% betavec[-j]
      betavec[j] <- sfun(sum(w*(dat$X[,j])*(dat$y - ynoj)), lambda)/(sum(w*(dat$X[,j])^2))
    }
    u <- dat$X %*% betavec
    expu <- exp(u) 
    loglik <- t(u) %*% dat$y - sum(t(log(1 + expu))) - sum(lambda * abs(betavec))
    # quardratic approximation to the log likelihood
    # loglik <- 1/2*(1/n)*sum(w*(z-u)^2) - sum(lambda * abs(betavec))
    res <- rbind(res, c(i, loglik, betavec))}  
  return(res)
}
corres <- coordinatelasso(lambda = 0.5, newdat, start = rep(0, dim(newdat$X)[2]))
check <- tail(corres)[,1:2]
cor.beta <- corres[nrow(corres),3:dim(corres)[2]]
```

## Results
### Eestimation path for Lasso

```{r}
# impletement the pathwise coordinatewise optimization algorithm to obtain a path of solutions
path <- function(inputx,inputy,grid){
  start <- rep(0, dim(inputx)[2])
  betas <- NULL
  for (x in 1:100) {
  cv.errors <- vector()
    cor.result <- coordinatelasso(lambda = grid[x], 
                                  dat = list(X=as.matrix(inputx),y=inputy),
                                  start= start)
    lasbeta <- cor.result[nrow(cor.result),3:dim(cor.result)[2]]
    start <- lasbeta
    betas <- rbind(betas,c(lasbeta))
  }
  return(data.frame(cbind(grid,betas)))
}
path.out <- path(newX,resp,grid=seq(0.5, 1e-2, length=100))
colnames(path.out) <- c("grid",colnames(newdataX),"the intercept")
# plot a path of solutions
path.plot <- path.out %>%
  gather(key = par, value = estimate, c(2:dim(path.out)[2])) %>% 
  ggplot(aes(x = grid, y = estimate, group = par, col = par)) +
  geom_line()+
  ggtitle("Figure 2: a path of solutions with a sequence of descending lambda's")+
  xlab("Lambda") + ylab("Estimate") 
path.plot
```


### Cross validation for lasso
one plot for beta
one plot for lambda

```{r}
##### 5-fold cross-validation and pathwise coordinatewise optimization algorithm
# when lambda = 0.5, all the betas go to 0
cvresult <- function(inputx,inputy,grid,K){
  n <- dim(inputx)[1]
  folds <- sample(1:K, n, replace=TRUE)
  start <- rep(0, dim(inputx)[2])
  cv.error <- vector()
  cv.se <- vector()
  for (x in 1:100) {
  cv.errors <- vector()
  for (i in 1:K){
    cor.result <- coordinatelasso(lambda = grid[x], 
                                  dat = list(X=as.matrix(inputx[folds!=i,]),y=inputy[folds!=i]),
                                  start= start)
    lasbeta <- cor.result[nrow(cor.result),3:dim(cor.result)[2]]
    u <- as.matrix(inputx[folds==i,])%*%lasbeta
    expu <- exp(u) 
    prob <- expu / (1 + expu) 
    pred <- ifelse(prob >=0.5,1,0)
    cv.errors[i] = mean((as.vector(inputy[folds==i])-as.vector(pred))^2) # MSE
    start <- lasbeta
  }
  cv.error[x] <- mean(cv.errors)
  cv.se[x] <- sqrt(var(cv.errors)/K)
  }
  return(cbind(grid,cv.error,cv.se))
}

result <- cvresult(newX,resp,grid=seq(0.5, 1e-2, length=100),K=5)
# result <- cvresult(X,resp,grid=seq(0.5, 1e-2, length=100),K=5)
best.lambda <- result[which.min(result[,2]),1]
finlasso <- coordinatelasso(lambda = best.lambda, newdat, start = rep(0, dim(newdat$X)[2]))
check <- tail(finlasso)[,1:2]
lasso.beta <- finlasso[nrow(finlasso),3:dim(finlasso)[2]]

# plot for cross validation
result <- data.frame(result)
cv.plot <- 
    ggplot(result, aes(x=result$grid, y=result$cv.error)) + 
    geom_errorbar(aes(ymin=result$cv.error-result$cv.se, ymax=result$cv.error+result$cv.se),
                  colour=1) +
    geom_line() +
    geom_point(size=1,colour = 4) +
    ggtitle("Figure 3: Lasso regression by 5 fold cross validation")+
    xlab("Lambda") + ylab("CV error") +
    geom_vline(xintercept = best.lambda,col=3,lty=3)
cv.plot
```


### Model comparison
comparison table 
```{r}
######## compare prediction performance of all results
pred.fun <- function(outcome,input, beta){
    u <- as.matrix(input)%*%beta
    expu <- exp(u) 
    prob <- expu / (1 + expu) 
    pred <- ifelse(prob >=0.5,1,0)
    pred.error = mean((as.vector(outcome)-as.vector(pred))^2)
    return(pred.error)
}
# logistic regression by GLM
log.beta <- c(logit.beta[2:length(logit.beta)],logit.beta[1])
log.pred <- mean((resp-predict(logfit.1,type="response"))^2) # abs(mean(logfit.1$residuals))
# newton's method
newton.ite <- nrow(newres)
newton.beta <- newres[nrow(newres),3:dim(newres)[2]]
newton.pred <- pred.fun(resp,newX,newton.beta)
# gradient decent
grad.ite <- nrow(gradres)
grad.beta <- gradres[nrow(gradres),3:dim(gradres)[2]]
grad.pred <- pred.fun(resp,newX,grad.beta)
# lasso logistic
lasso.ite <- nrow(finlasso)
lasso.beta <- finlasso[nrow(finlasso),3:dim(finlasso)[2]]
lasso.pred <-  pred.fun(resp,newX,lasso.beta)

beta.res <- round(as.matrix(rbind(log.beta,newton.beta,grad.beta,lasso.beta)),2)
colnames(beta.res) <- colnames(newX)
rownames(beta.res) <- c("GLM package","Newton Raphson","Gradient Decent","Logistic Lasso")
perf.res <- matrix(rep(NA),ncol = 2, nrow = 4)
colnames(perf.res) <- c("iteration times","prediction error")
rownames(perf.res) <- c("GLM package","Newton Raphson","Gradient Decent","Logistic Lasso")
perf.res[1,1] <- "NA"
perf.res[1,2] <- round(log.pred ,2)
perf.res[2,1] <- newton.ite
perf.res[2,2] <- round(newton.pred,2)
perf.res[3,1] <- grad.ite
perf.res[3,2] <- round(grad.pred,2)
perf.res[4,1] <- lasso.ite
perf.res[4,2] <- round(lasso.pred,2)

# output-performace
kable(t(perf.res), "latex", caption = "The comparison of performance for estimation algorithms and models", booktabs = T) %>% 
  kable_styling(latex_options = c("hold_position", "scale_down")) %>%
  add_footnote(c("Dataset: Breast Cancer Diagnosis"),
               notation = "alphabet") 
 
# output-beta
kable(t(beta.res), "latex", caption = "The comparison of performance for estimation algorithms and models", booktabs = T) %>% 
  kable_styling(latex_options = c("hold_position", "scale_down")) %>%
  add_footnote(c("Dataset: Breast Cancer Diagnosis"),
               notation = "alphabet")               

```






## Conclusions


## References
\begin{enumerate}
\item[1] Bender, Ralf, Thomas Augustin, and Maria Blettner. "Generating survival times to simulate Cox proportional hazards models." Statistics in medicine 24.11 (2005): 1713-1723.
\item[2] Austin, Peter C. "Generating survival times to simulate Cox proportional hazards models with time‐varying covariates." Statistics in medicine 31.29 (2012): 3946-3958.
\item[3] Morina, David, and Albert Navarro. "The R package survsim for the simulation of simple and complex survival data." Journal of Statistical Software 59.2 (2014): 1-20.
\end{enumerate}


## Appendix A
```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE}
```