---
title: "Group Projects on Optimization Algorithms."
date: "P8160 Advanced Statistical Computing "
output: pdf_document #html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(Hmisc)
set.seed(99999)
```

# Breast Cancer Diagnosis

The data \textit{breast-cancer.csv} have 569 row and 33 columns. The first column \textbf{ID} lables individual breast tissue images; The second column \textbf{Diagnonsis} indentifies if the image is coming from cancer tissue or benign cases (M=malignant, B = benign). There are 357  benign and 212  malignant cases. The other 30 columns correspond to mean, standard deviation and the largest values (points on the tails) of the distributions of the following 10 features computed for the cellnuclei;
\begin{itemize}
\item radius (mean of distances from center to points on the perimeter)
\item texture (standard deviation of gray-scale values)
\item perimeter
\item area
\item smoothness (local variation in radius lengths)
\item compactness (perimeter\^ 2 / area - 1.0)
\item concavity (severity of concave portions of the contour)
\item concave points (number of concave portions of the contour)
\item symmetry
\item fractal dimension ("coastline approximation" - 1)
\end{itemize}

The goal of the exerise is to build a predictive model based on logistic regression to facilicate cancer diagnosis; 

\textbf{Your to-do-list is}
```{r}
######## descriptive data analysis -one plot for all random variables in report
######## model comparison -table-estimation for betas in report
######## cross-validation -one plot appendix
######## model diagnostic -one plot appendix
```

\begin{enumerate}
\item Build a logistic model to classify the images into malignant/benign, and write down your likelihood function, its gradient and Hessian matrix.  
Let $y$ be the vector $n$ response random variable, $X$ denote the $n\times p$ design matrix(let$X_{i}$ denote the $i$th row) and $\beta$ denote the $p\times 1$ coefficient.
The likelihood of logistic regression is:
$$L(\beta; X, y) = \prod_{i=1}^n \{(\frac{\exp(X_{i}\beta)}{1+\exp(X_{i}\beta)})^{y_i}(\frac{1}{1+\exp(X_{i}\beta)})^{1-y_i}\}$$
Maximizing the likelihood is equivalent to maximizing the log likelihood:
$$
\begin{aligned}
f(\beta) 
& = \sum_{i=1}^n \{y_i(X_{i}\beta)-\log(1+\exp(X_{i}\beta))\}\\
& = <X\beta, Y> - \sum_{i=1}^n\log(1+\exp(X_{i}\beta))
\end{aligned}
$$
Let $p$, a vector of $n$ denote $p=\frac{\exp(X\beta)}{1+\exp(X\beta)}$. 
The gradient of this function is:
$$\nabla f(\beta) = X^T(y-p)$$
The Hessian is given by:
$$\nabla^2 f(\beta) = -X^T p(1-p)^TX$$
Hessian matrix is negative definite, well behaved.


```{r}
###### data manipulation
# setwd("/Users/yujingyao/Desktop/")
mydata <- read.csv("breast-cancer-1.csv") 
n <- dim(mydata)[1]
p <- dim(mydata)[2]
list <- c(3:(p - 1))
data <- do.call(cbind, lapply(list, function(x) (mydata[,x] - mean(mydata[,x]))/sd(mydata[,x]))) # standardize
X <- data.frame(data) %>% 
  mutate(., intercept = 1)
y <- as.vector(ifelse(mydata[,2] == "M", 1, 0))
newdat <- cbind.data.frame(y,X)
logfit <- glm(y~., family = binomial(link = "logit"),
          data = newdat)
summary(logfit)
# didn't converge
```

\item Develop a Newton-Raphson algoirthm to estimate your model; 

```{r}
##### 1. classical newton raphson
dat <- list(y = y, X = as.matrix(X))
logisticstuff <- function(dat, betavec){ 
  u <- dat$X %*% betavec
  expu <- exp(u) 
  loglik <- t(u) %*% dat$y - sum(t(log(1 + expu))) # Log-likelihood at betavec
  prob <- expu / (1 + expu) # P(Y_i=1|x_i), the p vector we mentioned before
  grad <- t(dat$X) %*% (dat$y - prob)
  Hess <- -t(dat$X) %*% prob %*% t(1 - prob) %*% dat$X # Hessian at betavec
  return(list(loglik = loglik, grad = grad, Hess = Hess))}

NewtonRaphson <- function(dat, func, start, tol=1e-10, maxiter = 200){ 
  i <- 0 
  cur <- start 
  stuff <- func(dat, cur) 
  res <- c(0, stuff$loglik, cur)
  prevloglik <- -Inf # To make sure it iterates 
  
  while (i < maxiter && abs(stuff$loglik - prevloglik) > tol && stuff$loglik > -Inf) {
    i <- i + 1 
    prevloglik <- stuff$loglik 
    prev <- cur # theta i
    cur <- prev - solve(stuff$Hess) %*% stuff$grad  # theta i+1
    stuff <- func(dat, cur) # log-lik, gradient, Hessian 
    res <- rbind(res, c(i, stuff$loglik, cur))}  
  
  return(round(res[nrow(res),3:ncol(res)], 2))
}
# ans1 <- NewtonRaphson(dat,logisticstuff, rep(1, 31), maxiter = 500)
# Error in solve.default(stuff$Hess) : singular
```

```{r}
##### 2. gradient descent
gradient <- function(dat, func, start, tol=1e-5, maxiter = 200){ 
  i <- 0 
  cur <- start 
  pp <- length(start)
  stuff <- func(dat, cur) 
  res <- c(0, stuff$loglik,cur)
  prevloglik <- -Inf # To make sure it iterates 
  while (i <= maxiter && abs(stuff$loglik - prevloglik) > tol && stuff$loglik > -Inf) { 
    i <- i + 1 
    prevloglik <- stuff$loglik 
    prev <- cur 
    cur <- prev + (diag(pp)/10) %*% (stuff$grad) # smaller step
    stuff <- func(dat, cur) # log-lik, gradient, Hessian 
    res <- rbind(res, c(i, stuff$loglik, cur))} 
  
  return(round(res[nrow(res),3:ncol(res)],2))
}
ans2 <- gradient(dat, logisticstuff, rep(0,31), maxiter = 500)
```

```{r}
##### 3. coordinate-wise decent update
coordinate <- function(dat, start, tol=1e-10, maxiter = 200) { 
  i <- 0 
  pp <- length(start)
  n <- length(dat$y)
  betavec <- start
  u <- dat$X %*% betavec
  expu <- exp(u) 
  loglik <- t(u) %*% dat$y - t(log(1 + expu)) %*% dat$y 
  res <- c(0, loglik, betavec)
  prevloglik <- -Inf # To make sure it iterates 
  while (i < maxiter && abs(loglik - prevloglik) > tol && loglik > -Inf) { 
    i <- i + 1 
    prevloglik <- loglik
    prev <- betavec
    for (j in 1:pp) {
      ynoj <- dat$X[,-j] %*% betavec[-j]
      betavec[j] <- 1/n * (t(dat$X[,j]) %*% (dat$y - ynoj))
    }
    u <- dat$X %*% betavec
    expu <- exp(u) 
    loglik <- t(u) %*% dat$y - t(log(1 + expu)) %*% dat$y 
    res <- rbind(res, c(i, loglik, betavec))}  
  
  return(round(res[nrow(res),3:ncol(res)],2))
}
ans3 <- coordinate(dat,rep(1,31))
```


\item Build a logistic-LASSO model to select features, and impletement the pathwise coordinatewise optimization algorithm to obtain a path of solutions with a sequence of descending $\lambda$'s. 
```{r}
##### 4.coordinate-wise decent update with lasso
sfun <- function(beta,lambda) sign(beta)*max((beta-lambda),0)
coordinatelasso <- function(lambda,dat, start, tol=1e-10,maxiter = 200){ 
  i <- 0 
  pp <- length(start)
  n <- length(dat$y)
  betavec <- start 
  u <- (dat$X)%*%betavec
  expu <- exp(u) 
  loglik <- t(u)%*%dat$y-t(log(1+expu))%*%dat$y + sum(lambda*abs(betavec))
  res <- c(0, loglik, betavec)
  prevloglik <- -Inf # To make sure it iterates 
  while(i < maxiter && abs(loglik - prevloglik) > tol && loglik > -Inf) 
    { i <- i + 1 
    prevloglik <- loglik
    prev <- betavec
    for (j in 1:pp){
      ynoj <- dat$X[,-j]%*%betavec[-j]
      betavec[j] <- sfun(1/n*(t(dat$X[,j])%*%(dat$y-ynoj)),lambda)
    }
    u <- dat$X%*%betavec
    expu <- exp(u) 
    loglik <- t(u)%*%dat$y-t(log(1+expu))%*%dat$y + sum(lambda*abs(betavec))
    res <- rbind(res, c(i, loglik, betavec))}  
  return(round(res[nrow(res),3:ncol(res)],2))
}
ans4 <- coordinatelasso(lambda =1e-2, dat,rep(1,31))
```

\item Use 5-fold cross-validation to select the best $\lambda$. Compare the prediction performance between the 'optimal' model and 'full' model  
```{r}
##### 5-fold cross-validation
set.seed (99999)
K <- 5
folds <- sample(1:K, n, replace=TRUE)
grid <- 10^seq(-2,0, length=100)
list <- seq(1:100)
cvresult <- do.call(cbind,lapply(list,function(x){
  cv.errors <- vector()
  for (i in 1:K){
  loglasbeta <- coordinatelasso(grid[x], 
                                list(X=as.matrix(X[folds!=i,]),y=y[folds!=i]),
                                start= rep(1,31),tol=1e-1,maxiter = 20)
  u <- as.matrix(X[folds==i,])%*%loglasbeta
  expu <- exp(u) 
  prob <- expu / (1 + expu) 
  pred <- ifelse(prob >=0.5,1,0)
  cv.errors[i] = mean((as.vector(y[folds==i])-as.vector(pred))^2) 
  }
  cv.errors.mean <- mean(cv.errors)
  cv.errors.se <- sqrt(var(cv.errors)/K)
  return(c(cv.errors.mean,cv.errors.se))
  }
))
result <- t(rbind(grid, cvresult))
colnames(result) <-c("grid","cv.error","cv.se")
best.lambda <- result[which.min(result[,2]),1]
final <- coordinatelasso(lambda =best.lambda, dat,rep(1,31))

# plot
result <- data.frame(result)
plot(result$grid, result$cv.error, type="n", xlab="Lambda",cex.lab=1,xaxt="n",cex.axis=1, cex.main=1,
     ylab="CV error", main="Figure 3: Lasso regression by 5 fold cross validation", ylim=c(0,1))
lines(result$grid, result$cv.error ,pch=16, col=4,lwd=1.5, type="b", cex=1)
with(data = result, expr = errbar(grid, cv.error, cv.error+cv.se, cv.error-cv.se, add=T, pch=1, cap=.015, log="x"))
abline(v=best.lambda,col=3,lty=3)
axis(side=1,at=c(0.2, 0.4, 0.6, 0.8, 1.0),labels=c("0.2", "0.4", "0.6","0.8",'1'), cex.axis=1)
```

\item Write a report to summarize your findings.
\end{enumerate}
# Answer: 
```{r }

```

