---
title: "A study of optimization algorithms on a breast cancer diagnosis dataset"
author: "Xinlei Chen, Guojing Wu and Yujing Yao"
abstract: "This report discusses a study of optimization algorithms on breast cancer diagnosis dataset. Our goal is to build a predictive model based on logistic regression to facilicate cancer diagnosis, and we compared estimation method including "
thanks:
keywords: "Newton "
date: "March 14, 2019"
output:
    pdf_document:
    highlight: default
    number_sections: true
    citation_package:
    keep_tex: false
    fig_caption: true
    latex_engine: pdflatex
fontsize: 10pt
geometry: margin=1in
bibliography:
biblio-style:
header-includes:
- \usepackage{indentfirst}
- \usepackage{graphicx}
- \usepackage{geometry}
- \usepackage{subfigure}
- \usepackage{amsmath}
- \usepackage{listings}
- \usepackage{tikz}
- \usetikzlibrary{matrix}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, comment = "")
require(dplyr) # dataset manipulation
require(plyr) # apply function
require(knitr) # table
require(kableExtra) # table
require(ggplot2) # for plot
require(ggpubr) # common legend
options(knitr.table.format = "latex")
```

## Background

1. Dataset descriptive
2. check multicolinearity -one plot
3. report final input of model - descripive mean(sd)


## Method

### Logistic Regression
problem 1

### Newton Raphson
problem2 

### Lasso
problem 3&4 

## Results

### Cross validation for lasso
one plot for beta
one plot for lambda

### Model comparison
comparison table 

\begin{table}[ht]
\caption{Data generation function for the exponential, Weibull, and Gompertz distributions.}
\begin{center}
\begin{tabular}{|l|c|c|c|}
\hline
Characteristic & exponential distribution & Weibull distribution &Gompertz distribution \\[1mm]\hline
Parameter & Scale parameter & Scale parameter $\lambda >0$& Scale parameter $\lambda >0$\\
 & $\lambda >0$ & Shape parameter $\gamma >0$& Shape parameter $\alpha \in R$\\[2mm]
Hazard function& $h_0(t) = \lambda$ & $h_0(t) = \lambda\gamma t^{\gamma-1}$ & $h_0(t) = \lambda e^{\alpha t}$\\[2mm]
Cum. hazard function & $H_0(t) = \lambda t$ & $H_0(t) = \lambda t^\gamma$ & $H_0(t) = \frac{\lambda}{\alpha}(e^{\alpha t} - 1)$
\\[2mm]
Survival time $u \sim U (0,1)$ & 
$T  = \frac{-logu}{\lambda \exp(\beta x)}$ & 
$T = (\frac{-logu}{\lambda \exp(\beta x)})^\frac{1}{\gamma}$ & 
$T = \frac{1}{\alpha} log(1 - \frac{\alpha logu}{\lambda \exp(\beta x)})$
\\[1mm]\hline
\end{tabular}
\end{center}
\label{default}
\end{table}


**Figure 1: Trajectory plot of estimation criteria($\beta=1.5$, $\lambda = 0.1$, $\alpha=5$, $\nu=5$, non censored).**

**Figure 2: Trajectory plot of estimation criteria($n=300$, $\lambda = 0.1$, $\alpha=5$, $\nu=5$, non censored).**



## Conclusions


## References
\begin{enumerate}
\item[1] Bender, Ralf, Thomas Augustin, and Maria Blettner. "Generating survival times to simulate Cox proportional hazards models." Statistics in medicine 24.11 (2005): 1713-1723.
\item[2] Austin, Peter C. "Generating survival times to simulate Cox proportional hazards models with time‐varying covariates." Statistics in medicine 31.29 (2012): 3946-3958.
\item[3] Morina, David, and Albert Navarro. "The R package survsim for the simulation of simple and complex survival data." Journal of Statistical Software 59.2 (2014): 1-20.
\end{enumerate}

## Appendix A



## Appendix B
```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE}
```